---
title: "Exponential Distribution and CLT"
output: html_document
---


# Part 1 Exponetial and CLT

## Overview

#### In this project, I will investigate the exponential distribution in R and compare it with the Central Limit Theorem: the distribution of averages of 40 exponentials with a thousand simulations.

## Show the sample mean and compare it to the theoretical mean of the distribution.

#### Theortical Mean

```{r}
lambda <- 0.2
T_mean <- 1/lambda
print(T_mean)
```



#### Sample Mean
```{r}
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40,lambda)))
S_mean<-mean(mns)
print(S_mean)
head(mns)
```


#### compare distribution
```{r}
hist(mns, main="The distribution of 1000 averges of 40 random exponetials", xlab="Value of means", ylab="Frequecy of means", col="blue")
abline(v=T_mean, lty=1, lwd=5, col="red")
legend("topright", lty=1, lwd=5, col="red", legend="theoretical mean")
```

## Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

#### Theortical variance
```{r}
T_variance <- (1/lambda)^2
print(T_variance)
```

#### Sample variance
```{r}
vns = NULL
for (i in 1 : 1000) vns = c(vns, var(rexp(40,lambda)))
S_variance <- var(vns)
print(S_variance)
head(vns)
```

#### compare distribution
```{r}
hist(vns, main="The distribution of 1000 variance of 40 random exponetials", xlab="Value of variances", ylab="Frequecy of variances", col="blue")
abline(v=T_variance, lty=1, lwd=5, col="red")
legend("topright", lty=1, lwd=5, col="red", legend="theoretical variance")
```



## Show that the distribution is approximately normal:focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

```{r}
ns = NULL
for (i in 1 : 1000) ns = c(ns, rexp(40,lambda))

par(mfrow = c(3, 1))
hist(ns, main = "Distribution of exponentials with lambda equals to 0.2", xlab = "Exponentials", col = "yellow")
hist(mns, main = "The distribution of 1000 averages of 40 random exponentials", xlab = "Value of means", ylab = "Frequency of means", col = "blue")
Norm <- rnorm(1000, mean = mean(mns), sd = sd(mns))
hist(Norm, main = "A normal distribution with theoretical mean and sd of the exponentials", xlab = "Normal variables", col = "red")
```

#### summary: as a result of the central limit theorem, comparing the second and the third histogram, we can see the distribution of the means is similar to a real normal distribution with the same mean and standard deviation.


# Part 2 ToothGrowth Analysis

## Overview : analyze tooth growth data and have some conclusions

#### load data/libarary, basic summary and explore data
```{r}
library(ggplot2)
data("ToothGrowth")
dim(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```

#### plot and see the trend
```{r}
ggplot(ToothGrowth, aes(x = dose, y=len, color=supp)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title = "ToothGrowth trend based on supplements", x = "Dose of supplement ", y = "Length of teeth")
```

#### initial conclusions:

* supplment and tooth growth are linear relationship
* same dose, OJ has bigger effect compared with VC
* VC has larger gradient of increase compared with OJ

#### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose
#### assumption -- normal distribution, null hypothesis -- no difference between mean of 2 kinds of supplements or dose
```{r}
t.test(x = ToothGrowth$len, data = ToothGrowth, paired = FALSE, conf.level = 0.95)$conf.
```
#### calculate confidence level and p value based on kind of supplement
```{r}
t.test(ToothGrowth[ToothGrowth$supp == "OJ", ]$len)$p.value
t.test(ToothGrowth[ToothGrowth$supp == "OJ", ]$len)$conf.int
t.test(ToothGrowth[ToothGrowth$supp == "VC", ]$len)$p.value
t.test(ToothGrowth[ToothGrowth$supp == "VC", ]$len)$conf.int
```

####  calculate confidence level and p value based on dose
```{r}
t.test(ToothGrowth[ToothGrowth$dose == 0.5, ]$len)$p.value
t.test(ToothGrowth[ToothGrowth$dose == 0.5, ]$len)$conf.int
t.test(ToothGrowth[ToothGrowth$dose == 1, ]$len)$p.value
t.test(ToothGrowth[ToothGrowth$dose == 1, ]$len)$conf.int
t.test(ToothGrowth[ToothGrowth$dose == 2, ]$len)$p.value
t.test(ToothGrowth[ToothGrowth$dose == 2, ]$len)$conf.int
```


#### calculate mean based on dose for each supplement

```{r}
dose05 <-ToothGrowth[ToothGrowth$dose == 0.5, ]
mean(dose05[dose05$supp=="OJ",]$len)
mean(dose05[dose05$supp=="VC",]$len)
```

```{r}
dose1 <-ToothGrowth[ToothGrowth$dose == 1, ]
mean(dose1[dose1$supp=="OJ",]$len)
mean(dose1[dose1$supp=="VC",]$len)
```

```{r}
dose2 <-ToothGrowth[ToothGrowth$dose == 2, ]
mean(dose2[dose1$supp=="OJ",]$len)
mean(dose2[dose1$supp=="VC",]$len)
```

#### OJ delivers more tooth growht than VC for dose 0.5 & 1. OJ and VC deliver same tooth growth for dose 2.

